Contracts for unions and intersections are hard for Nickel

### 2021-05-16 paper.tex
Tiny review on flat contracts

@@

For instance, we could define a flat contract $even?$ as
$\lambda x. ~ x ~ \% ~ 2 ~ == ~ 0$, that checks
whether a number is even by checking the remainder of dividing it by 2.
And then use it to check that a function always returns an
even number, when applied to a $Number$:

$$let~f~=~(\lambda n. ~ n * 2)@(Number ~ \rightarrow ~ even?)$$

@@

This kind of contracts are present on many different languages,
for instance, the Eiffel programming language, the precursor
of the Design by Contract ideals, allows to assert
these kinds of expression as pre- and post-conditions on
functions\cite{EiffelDesignByContract}.
A somewhat newer language, the D programming
language, allows, among other things, to have invariants on classes,
written with flat contracts\cite{DLangContracts}.

Finally, the Racket programming language also has a system to work with
contracts\footnote{And a gradual typing system
built on top of that, but that goes beyond the scope of this
work.} that allows to define flat contracts, and
to compose them with other kinds of dynamic checks,
like higher order contracts or a lightweight take on union
and intersection contracts\cite{RacketContracts}.
\todo{Talk about this, probably later on the paper}

Even if flat contracts may seem like too much, the ability to make
non trivial checks on values (to validate their output) is expected
by programmers on dynamic languages, where testing and runtime
assertions are a common pattern for safe programming, and
programmers are used to express complex predicates
over values.

@@

\section{Related work}
\info{Including a mention of statically typed systems with union}

>>>>

\newcommand{\nickel}[1]{\lstinline[language=nickel]{#1}}

@@

For instance, we could define a flat contract \nickel{even} as
\nickel{fun x => x \% 2 == 0}, that checks
whether a number is even by checking the remainder of dividing it by 2.
And then use it to check that a function always returns an
even number, when applied to a \nickel{Number}:

\begin{figure}[h]
\begin{center}
\begin{lstlisting}[language=nickel]
let f = (fun n => n * 2) @ (Number -> even)
\end{lstlisting}
  \end{center}
\caption{A function wrapped with a contract stating that it always returns
an even number}
\label{fig:num-to-even}
\end{figure}

% $$let~f~=~(\lambda n. ~ n * 2)@(Number ~ \rightarrow ~ even?)$$

@@

This kind of contracts are present on many different languages,
for instance, the Eiffel programming language, the precursor
of the Design by Contract ideals, allows to assert
these kinds of expression as pre- and post-conditions on
functions\cite{EiffelDesignByContract}.
A somewhat newer language, the D programming
language, allows, among other things, to have invariants on classes,
written with flat contracts, that are
dinamically checked at the frontiers of every public method \cite{DLangContracts}.

Finally, the Racket programming language also has a system to work with
contracts that allows to define flat contracts
\footnote{And a gradual typing system
built on top of that, but that goes beyond the scope of this
work.}
, and
to compose them with other kinds of dynamic checks,
like higher order contracts or a lightweight take on union
and intersection contracts\cite{RacketContracts}.
\todo{Talk about this, probably later on the paper.
See Related Works section}

Even if flat contracts may seem like too permissive, the ability to make
non trivial checks on values (to validate their output) is expected
by programmers on dynamic languages, where testing and runtime
assertions are a common pattern for safe programming, and
programmers are used to express complex predicates
over values.

@@

\section{Related work}
\info{Including a mention of statically typed systems with union}
\info{Include something about Racket's and and or contracts}
